# Reproducible Research-Project 1

# R preparations
In this document code will be represented to show how the results have been achieved. Set the default of echo to be true throughout the document:

# Load in the necessary packages
```{r}
library(knitr)
opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(ggplot2)
```

# Loading and preprocessing the data
The data is loaded using the read.csv()

# Reading in the data
```{r}
dataSet <- read.csv("activity.csv", header = TRUE, sep = ',')
```


# What is mean total number of steps taken per day?
1. Calculate the total number of steps per day:

```{r}
by_day<-aggregate(steps~date,dataSet,sum)
by_day$date<-as.Date(by_day$date,format="%Y-%m-%d")
```

2. Use ggplot for making the histogram:

```{r}
ggplot(by_day, aes(x=date, y=steps)) + geom_bar(stat="identity") +
  #scale_x_datetime(breaks="1 week", labels=date_format("%d-%m")) +
  ylab("steps") + xlab("Day and Month 2012") +ggtitle("Steps each day")
```

3. Calculate the mean and median of the total number of steps taken per day:

```{r}
by_dayMean<-mean(by_day$steps,na.rm=TRUE)
by_dayMedian<-median(by_day$steps,na.rm=TRUE)
by_dayMean
by_dayMedian
```


# What is the average daily activity pattern?
1. Calculate the average number of steps taken in each 5-minute interval per day using dplyr and group by interval:

```{r}
by_interval<-aggregate(steps~interval,dataSet,mean)
head(by_interval)
```

Use ggplot for making the time series of the 5-minute interval and average steps taken:

```{r}
ggplot(by_interval, aes(x=interval, y=steps)) +
  geom_line(color = "firebrick") +ggtitle("Average activity pattern")
```

2. Use which.max() to find out the maximum steps, on average, across all the days:

```{r}
by_interval[max(by_interval$steps),]
```


# Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.


1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NA):
```{r}
sum(is.na(dataSet$steps))
```

2.Devise a strategy for filling in all of the missing values in the dataset with the average of the steps in the others days. 
```{r}
data_full <- dataSet
j<-0
for (i in 1:nrow(data_full)){
  j<-j+1
  if (is.na(data_full[i,1])){
    data_full[i,1]<-by_interval[j,2]
  }
  
  if(data_full[i,2]!=data_full[i+1,2] & i+1<nrow(data_full))
    j<-0
}
```


3.Create a new dataset as the original and use tapply for filling in the missing        values with the average number of steps per 5-minute interval:
```{r}
head(data_full)
```


4. Calculate the number of steps taken in each 5-minute interval per day using dplyr     and group by interval. Use ggplot for making the histogram:
```{r}
by_day_Integrate<-aggregate(steps~date,data_full,sum)
by_day_Integrate$date<-as.Date(by_day_Integrate$date,format="%Y-%m-%d")
```


```{r}
ggplot(by_day_Integrate, aes(x=date, y=steps)) + geom_bar(stat="identity") +
  #scale_x_date(breaks="1 week", labels=date_format("%d-%m")) +
  ylab("steps") + xlab("Day and Month 2012") +ggtitle("Steps each day")
```

# Calculate the mean and median steps with the filled in values:
```{r}
by_day_Integrate_Mean<-mean(by_day_Integrate$steps,na.rm=TRUE)
by_day_Integrate_Median<-median(by_day_Integrate$steps,na.rm=TRUE)
by_day_Integrate_Mean
by_day_Integrate_Median
```



# Are there differences in activity patterns between weekdays and weekends?
For this part the weekdays() will come handy. Use the dataset with the filled-in missing values for this part.

1. Create a new factor variable in the dataset with two levels - “weekday” and           “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
data_full$date<-as.Date(data_full$date,format="%Y-%m-%d")

for (i in 1:nrow(data_full)){
  if (weekdays(data_full[i,2])=="samedi" | weekdays(data_full[i,2])=="dimanche"){
    data_full[i,4]<-"Weekend"
  }
  else
    data_full[i,4]<-"Weekday"
}

```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute     interval (x-axis) and the average number of steps taken, averaged across all          weekday days or weekend days (y-axis).
```{r}
interval_full <- data_full %>%
  group_by(interval, V4) %>%
  summarise(steps = mean(steps))

ggplot(interval_full, aes(x=interval, y=steps, color = V4)) +
  geom_line() +ggtitle(("differences in activity between weekdays and weekends"))+
  facet_wrap(~V4, ncol = 1, nrow=2)

```